### 2004-06-30 Documentation/part1.tex
*** empty log message ***

\example{Using \GLOBES\ with C}{\label{ex:c}
\index{C-Code}
Here comes the C-code sceleton, which is
(more or less) common to all of our \GLOBES\ examples:
\begin{quote}
{\tt {\footnotesize
\#include <stdio.h> \\
\#include <stdlib.h> \\
\#include <math.h> \\
\#include <string.h> \\
\\
\#include <globes/globes.h> \hspace*{0.5cm} /* Include GLoBES library */ \\
\\
\#include "myio.h" \hspace*{0.5cm} /* Include "housemade" I/O-routines */ \\
\\
/* If filename given, write to file; if empty, to screen: */ \\
char MYFILE[]="testX.dat"; \\
\\
int main(int argc, char *argv[]) \\
\{  \\
\\
 \hspace*{0.5cm} GLBInit(argv[0]); \hspace*{0.5cm} /* Initialize GLoBES library */ \\
\\  
  \hspace*{0.5cm} glbInitExperiment("NuFact.glb",\&glb\_experiment\_list[0], \\
  \hspace*{1cm} \&glb\_num\_of\_exps); \hspace*{0.5cm} /* Initialize experiment NuFact.glb */\\
\\  
  \hspace*{0.5cm}  /* Initialize housemade output function */\\
  \hspace*{0.5cm} 
   InitOutput(MYFILE,"Format: ... ... ... $\backslash$n"); \\
\\  
  \hspace*{0.5cm} /* Initialize parameter vector(s) */ \\
  \hspace*{0.5cm} glb\_params true\_values = glbAllocParams(); \\
  \hspace*{0.5cm} /* ... */ \\
\\  
 \hspace*{0.5cm} /* Assign: theta12,theta13,theta23,deltacp,dm2solar,dm2atm */ \\
   \hspace*{0.5cm}  
     glbDefineParams(true\_values,\\
     \hspace*{1.5cm}asin(sqrt(0.8))/2,asin(sqrt(0.001))/2,M\_PI/4,M\_PI/2,7e-5,2e-3); \\

@@

\example{Correlation between $\stheta$ and $\deltacp$}{
\label{ex:corrth13dcp}

\index{Two-parameter correlation}
A typical application for {\tt Chi} and {\tt SingleChi} is the visualization of two-parameter correlations. For example, to show the correlation between $\stheta$ and $\deltacp$ at a large neutrino factory with four years of running time for each polarity, one can use the following code:
\begin{quote}
{\tt
/\% Experiment initialization: \%/ \\
ClearExp(); \\
 InitExperiment("{\tt NuFact2.exp}"); \\
 SetRunningTime(0,4.0); \\
 MInfoExperiment(); \\
\\
/\% Set reference rate vector: \%/ \\
th12=arcsin(sqrt(0.8))/2; sdm=7e-5; \\
th23=3.14/4; ldm=2e-3; \\
th13=arcsin(sqrt(0.001))/2; dcp=3.14/2; \\
\mbox{SetVacuumParameters(\{th12, \, th13, \, th23, \, dcp, \, sdm, \, ldm\}); }\\
 SetRates(); \\
\\
/\% Compute chi-square matrix: \%/ \\
for (float x=-4; x<-2; x=x+2/50) \\
\hspace*{0.5cm} for (float y=0; y<200; y=y+200/50) \{ \\
\hspace*{1cm} theth13 = arcsin(sqrt(10$\hat{\, \, \,}$x))/2; \\
\mbox{\hspace*{1cm} Print(\{x,y,Chi(\{th12, theth13,  th23,  y, sdm, ldm, 1.0\}) \}) \};}
}
\end{quote}

The resulting matrix can then be plotted as a contour plot (2 d.o.f.):
\begin{center}
\colorbox{white}{\includegraphics[width=8cm]{correx}}
\end{center}

@@

This example demonstrates how to use  {\tt ChiTheta} to project the fit manifold onto the $\stheta$-axis, \ie, how one can include correlations. We compute two sets of data: one for keeping all parameters but $\deltacp$ fixed (two-parameter correlations), and one for keeping all parameters free (multi-parameter correlation), but some external input for the solar parameters and $\rho$. Note that fixing all parameters but $\deltacp$ corresponds to imposing external knowledge on these parameters.

\begin{quote}
{\tt {\footnotesize
/\% Experiment initialization, set reference rate vector: \%/ \\
\mbox{ClearExp(); InitExperiment("{\tt NuFact2.exp}"); SetRunningTime(0,4.0); }\\
th12=arcsin(sqrt(0.8))/2; sdm=7e-5; th23=3.14/4; ldm=2e-3; \\
th13=arcsin(sqrt(0.001))/2; dcp=3.14/2; \\
\mbox{SetVacuumParameters(\{th12, th13, th23, dcp, sdm, ldm\}); SetRates();}\\
\\
/\%  Compute chi-square list with all but dcp fixed: \%/ \\
SetStartingValues(\{th12, th13, th23, dcp, sdm, ldm, 1.0\}); \\
\mbox{SetInputErrors(\{th12*0.001, 10, th23*0.001, 10, sdm*0.001, ldm*0.001, 0.001\});} \\
for (float x=-4; x<-2; x=x+2/20) \\
\mbox{\hspace*{0.5cm} Print(\{x,ChiDelta(x,\{th12, th13, th23, dcp, sdm, ldm, 1.0\}) \}); }\\
\\
/\% Compute chi-square list with all parameters free: \%/ \\
\mbox{/\% (but: 10\% prec. for solar params, 5\% for matter density) \%/} \\
SetStartingValues(\{th12, th13, th23, dcp, sdm, ldm, 1.0\}); \\
SetInputErrors(\{th12*0.1, 10, 10, 10, sdm*0.1, ldm, 0.05\}); \\
for (float x=-4; x<-2; x=x+2/20) \\
\mbox{\hspace*{0.5cm} Print(\{x,ChiDelta(x,\{th12, th13, th23, dcp, sdm, ldm, 1.0\}) \});}
}}
\end{quote}
The two lists of data then represent the $\stheta$ precisions with two-parameter correlations (gray-shaded) and multi-parameter correlations (arrows):
\begin{center}
\colorbox{white}{\includegraphics[width=6cm]{projallex}}

@@

In many cases, one can find the $\mathrm{sgn}(\ldm)$-degeneracy with {\tt ChiAll}: One lets the local minimizer start at the position of the suspected solution at the opposite sign of $\ldm$.  The following example corresponds to finding the degenerate solution for the example on page~\pageref{ex:corrth13dcp}.
\begin{quote}
{\tt
{\footnotesize
/\% Experiment initialization: \%/ \\
ClearExp(); InitExperiment("{\tt NuFact2.exp}"); SetRunningTime(0,4.0); \\
\\
/\% Set reference rate vector: \%/ \\
th12=arcsin(sqrt(0.8))/2; sdm=7e-5; th23=3.14/4; ldm=2e-3; \\
th13=arcsin(sqrt(0.001))/2; dcp=3.14/2; \\
\mbox{SetVacuumParameters(\{th12, \, th13, \, th23, \, dcp, \, sdm, \, ldm\});  SetRates();} \\
\\
/\% Find negative-sign solution: \%/ \\
SetStartingValues(\{th12, \, th13, \, th23, \, dcp, \, sdm, \, -ldm, \, 1.0\}); \\
SetInputErrors(\{th12*0.1, \, 10, \, 10, \, 10, \, sdm*0.1, \, ldm/3, \, 0.05\}); \\
d = ChiAll(\{th12, \, th13, \, th23, \, dcp, \, sdm, \, ldm, \, 1.0\}); \\
\\
/\% Compute chi-square matrix: \%/ \\
for (float x=-4; x<-2; x=x+2/50) \\
\hspace*{0.5cm} for (float y=0; y<200; y=y+200/50) \{ \\
\hspace*{1cm} theth13 = arcsin(sqrt(10$\hat{\, \, \,}$x))/2; \\
\mbox{\hspace*{1cm} Print(\{x,y,Chi(\{d[2], theth13,  d[4],  y, d[6], d[7], d[8]\}) \}) \};}
}
}
\end{quote}

The resulting matrix can then be plotted as a contour plot in addition to the original solution (2 d.o.f., gray contours):
\begin{center}
\colorbox{white}{\includegraphics[width=7.5cm]{correntex}}
\end{center}

@@

\example{The impact of systematics, correlations, and degeneracies}{
\label{ex:corrproj}
\index{Bar charts}
\index{Systematics on/off}

Here it is shown how one can successively include systematics, correlations,
and degeneracies at the example of the $\stheta$ sensitivity limit.
An important part of this example is how two switch the systematics off,
\ie, how to obtain the sensitivity limit from statistics only. Since
this example is very advanced, we only show the respective part of the code:
\begin{quote}
{\tt 
Here comes the code excerpt.
}
\end{quote}
The complete code can be found as ``example4.c'' with the software.

The returned lists of data then represent $\Delta \chi^2$ 
as function of the fit value of $\stheta$. The intersections of these
curves with the line $\Delta \chi^2 = 9$ give the $\stheta$ sensitivity
limits, where we do not include the $\mathrm{sgn}(\ldm)$- and $(\deltacp,\theta_{13})$-degeneracies in the sensitivity limit with
correlations only (green bar):
\begin{center}
\colorbox{white}{\includegraphics[width=7cm]{barsex}}

\end{center}
}

>>>>

\example{Using \GLOBES\ with C}{\label{ex:c}
\index{C-Code}
Here comes the C-code sceleton, which is
(more or less) common to all of our \GLOBES\ examples:
\begin{quote}
{\tt {\footnotesize
\#include <stdio.h> \\
\#include <stdlib.h> \\
\#include <math.h> \\
\#include <string.h> \\
\\
\#include <globes/globes.h> \hspace*{0.5cm} /* Include GLoBES library */ \\
\\
\#include "myio.h" \hspace*{0.5cm} /* Include "housemade" I/O-routines */ \\
\\
/* If filename given, write to file; if empty, to screen: */ \\
char MYFILE[]="testX.dat"; \\
\\
int main(int argc, char *argv[]) \\
\{  \\
\\
 \hspace*{0.5cm} glbInit(argv[0]); \hspace*{0.5cm} /* Initialize GLoBES library */ \\
\\  
  \hspace*{0.5cm} glbInitExperiment("NuFact.glb",\&glb\_experiment\_list[0], \\
  \hspace*{1cm} \&glb\_num\_of\_exps); \hspace*{0.5cm} /* Initialize experiment NuFact.glb */\\
\\  
  \hspace*{0.5cm}  /* Initialize housemade output function */\\
  \hspace*{0.5cm} 
   InitOutput(MYFILE,"Format: ... ... ... $\backslash$n"); \\
\\  
  \hspace*{0.5cm} /* Initialize parameter vector(s) */ \\
  \hspace*{0.5cm} glb\_params true\_values = glbAllocParams(); \\
  \hspace*{0.5cm} /* ... */ \\
\\  
 \hspace*{0.5cm} /* Assign: theta12,theta13,theta23,deltacp,dm2solar,dm2atm */ \\
   \hspace*{0.5cm}  
     glbDefineParams(true\_values,\\
     \hspace*{1.5cm}asin(sqrt(0.8))/2,asin(sqrt(0.001))/2,M\_PI/4,M\_PI/2,7e-5,2e-3); \\

@@

\example{Correlation between $\stheta$ and $\deltacp$}{
\label{ex:corrth13dcp}
\index{Two-parameter correlation}
A typical and fast application for \GLB{glbChiSys} is the visualization of
 two-parameter correlations. For example, to calculate the two-parameter 
 correlation between $\stheta$ and $\deltacp$ at a neutrino factory, one
 can use the following code excerpt from {\tt example1.c}:
\begin{quote}
{\tt {\footnotesize
  /* Initialize parameter vector(s) and compute simulated data */ \\
  glbDefineParams(true\_values,theta12,theta13,theta23,deltacp,sdm,ldm); \\
  glbDefineParams(test\_values,theta12,theta13,theta23,deltacp,sdm,ldm);  \\
  glbSetOscillationParameters(true\_values); glbSetRates(); \\
\\
  /* Iteration over all values to be computed */ \\
  double x,y,res; \\   
  for(x=-4.0;x<-2.0+0.01;x=x+2.0/50) \\
  for(y=0.0;y<200.0+0.01;y=y+200.0/50) \\
  \{ \\
\hspace*{0.5cm} /* Set parameters in vector of test values */ \\
\hspace*{0.5cm} glbSetOscParams(test\_values,asin(sqrt(pow(10,x)))/2,GLB\_THETA\_13); \\
\hspace*{0.5cm} glbSetOscParams(test\_values,y*M\_PI/180.0,GLB\_DELTA\_CP); \\
  \\
\hspace*{0.5cm} /* Compute Chi2 for all loaded experiments and all rules */ \\
\hspace*{0.5cm} res=glbChiSys(test\_values,GLB\_ALL,GLB\_ALL); \\
\hspace*{0.5cm} AddToOutput(x,y,res); \\
  \} \\
} }
\end{quote}

The resulting data can then be plotted as a contour plot (2 d.o.f.):
\begin{center}
\colorbox{white}{\includegraphics[width=8cm]{correx}}
\end{center}

@@

This example demonstrates how to project the fit manifold onto the $\stheta$-axis, \ie, how one can include correlations. We compute two sets of data: one for keeping all parameters but $\deltacp$ fixed (two-parameter correlation), and one for keeping all parameters free (multi-parameter correlation). However, we impose external precisions for the solar parameters and the matter density. The following code excerpt is from 
{\tt example2.c}:

\begin{quote}
{\tt {\scriptsize
  /* Set starting values and input errors for all projections */  \\
  glbDefineParams(input\_errors,theta12*0.1,10,10,10,sdm*0.1,ldm);\\  
  glbSetDensityParams(input\_errors,0.05,GLB\_ALL);\\
  glbSetStartingValues(true\_values);\\
  glbSetInputErrors(input\_errors);\\
\\
  /* Define two-parameter projection: Only deltacp is free! */ \\
  glbDefineProjection(th13\_projection,GLB\_FIXED,GLB\_FIXED,GLB\_FIXED,GLB\_FREE,GLB\_FIXED,GLB\_FIXED); \\
  glbSetProjection(th13\_projection); \\ 
\\
  /* Iteration over all values to be computed */ \\
  double x,res1,res2;     \\
  for(x=-4;x<-2.0+0.001;x=x+2.0/50) \\
  \{ \\
\hspace*{0.5cm}  /* Set fit value of stheta */ \\
\hspace*{0.5cm} glbSetOscParams(test\_values,asin(sqrt(pow(10,x)))/2,1); \\
    \\
\hspace*{0.5cm} /* Guess fit value for deltacp in order to safely find minimum */ \\
\hspace*{0.5cm} glbSetOscParams(test\_values,200.0/2*(x+4)*M\_PI/180,3); \\
 \\
\hspace*{0.5cm} /* Compute Chi2 for user-defined two-parameter correlation */ \\
\hspace*{0.5cm} res1=glbChiNP(test\_values,NULL,GLB\_ALL); \\
      \\
\hspace*{0.5cm} /* Compute Chi2 for full correlation: minimize over all but theta13 */ \\
\hspace*{0.5cm} res2=glbChiTheta(test\_values,NULL,GLB\_ALL); \\
  \\
\hspace*{0.5cm} AddToOutput(x,res1,res2);\\
  \} \\

}}
\end{quote}
The two lists of data then represent the $\stheta$ precisions with two-parameter correlations (gray-shaded) and multi-parameter correlations (arrows):
\begin{center}
\colorbox{white}{\includegraphics[width=6cm]{projallex}}

@@

In many cases, one can find the exact position of the $\mathrm{sgn}(\ldm)$-degeneracy with \GLB{glbChiAll}, where one starts 
the local minimizer at the suspected position 
and let it converge into the minimum.  The following code excerpt  corresponds to finding the degenerate solution for the example on
  page~\pageref{ex:corrth13dcp}, and it is from {\tt example3.c}: 
\begin{quote}
{\tt
{\footnotesize
  /* Set starting vales to suspected position at opposite sign of ldm */ \\
  \mbox{glbDefineParams(starting\_values,theta12,theta13,theta23,deltacp,sdm,-ldm);} \\
  \\
  /* Set input errors for external input, where some information on ldm \\
   \hspace*{0.5cm} is imposed in order to avoid falling into the right-sign solution */ \\
  glbDefineParams(input\_errors,theta12*0.1,10,10,10,sdm*0.1,ldm/3); \\ 
  glbSetDensityParams(input\_errors,0.05,GLB\_ALL); \\
  \\
  glbSetStartingValues(starting\_values); \\
  glbSetInputErrors(input\_errors); \\
  \\
  /* Localize degenerate solution by minimzation over all parameters */ \\
  double CL=glbChiAll(starting\_values,deg\_pos,GLB\_ALL); \\
   \\
  \mbox{/* Now: CL is the chi2 of the deg.~solution and deg\_pos the position */} 
}}
\end{quote}

Using {\tt ent\_pos} to obtain a section of the degeneracy in the
$\stheta$-$\deltacp$-plane (\cf, {\tt example3.c}), one can plot it as a contour plot in addition to the original solution (2 d.o.f., gray curves):
\begin{center}
\colorbox{white}{\includegraphics[width=7.5cm]{correntex}}
\end{center}

@@

\example{The impact of systematics, correlations, and degeneracies}{
\label{ex:corrproj}
\index{Bar charts}
\index{Systematics on/off}
Here it is demonstrated how one can successively include systematics, correlations, and degeneracies at the example of the $\stheta$-sensitivity limit.
An important part of this example is how two switch the systematics off,
\ie, how to obtain the sensitivity limit from statistics only. Since
this example is very advanced, we only show the respective function of the code:
\begin{quote}
{\tt 
/* Calculate chi2 with statistics only */ \\
double CalcNoSystematics(double theldm,double thex) \\
\{ \\
\hspace*{0.5cm} /* Switch systematics off for all exps and all rules */ \\
\hspace*{0.5cm} glbSwitchSystematics(GLB\_ALL,GLB\_ALL,GLB\_OFF); \\
\\
\hspace*{0.5cm} /* Calculate Chi2-list as if systematics were on */ \\
\hspace*{0.5cm} double res=CalcSystematics(theldm,thex); \\
\\
\hspace*{0.5cm} /* Switch systematics on for all exps and all rules */ \\
\hspace*{0.5cm} glbSwitchSystematics(GLB\_ALL,GLB\_ALL,GLB\_ON); \\
\\
\hspace*{0.5cm} return res; \\
\} \\
}
\end{quote}

\vspace*{-0.5cm}

The complete code can be found as {\tt example4.c} with the software,
which consists of many of the applications from the earlier examples.
In addition, it uses a little trick: It avoids falling into the wrong
 solution with {\tt glbChiTheta}
by using the fit value of $\deltacp$ from the step before as prediction
of the position of the current calculation.

The returned lists of data from thie example represent $\chi^2$ 
as function of the fit value of $\stheta$. The intersections of these
curves with the line $\chi^2 = 9$ give the $\stheta$ sensitivity
limits at the $3 \sigma$ confidence level, where we do not include the
 $\mathrm{sgn}(\ldm)$- and $(\deltacp,\theta_{13})$-degeneracies in the sensitivity limit with correlations only (green bar):
\begin{center}
\colorbox{white}{\includegraphics[width=6.5cm]{barsex}}
\end{center}
}
